---
title: "**DrosEU Phenotyping Working Group Updates **"
author: "by Esra Durmaz Mitchell"
date: "Last updated: `r format(Sys.Date())`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cosmo  
---


# Phenotyping effort

General information  
How many labs  
How many people  


## Groningen 2017

- Fly husbandry @ IGC

- Phenotypes

- Decision to be made 

## Samples

- Locations

- Sampling 


```{r cars}

#Add samples map here, with people's names 

```


- Shipment to IGC

- Establishing lines @ IGC


## Phenotyping Call

- A few criteria for phenotyping 

- re : email to Elio about immunity 

- Responders

- Distribution of traits 

- Shipment of flies

- Phenotyping starts 

```{r}

#Add contributors map here 

```


# Phenotyping experiments 

## Overview 

List of traits in groups (JVieira's classification)

## Life-histroy traits 

### Viability

Sonja Grath (S), Marina Stamenkovic-Radak (S), Katja Hoedjes (NS), Patricia Gibert (NS)

### Developmental time

Sonja Grath (S), Marina Stamenkovic-Radak (S), Katja Hoedjes (NS), Patricia Gibert (NS)

### Fecundity

Eran Tauber (S), Claudia Fricke (S), Jean-Christophe Billeter (NS) 

### Lifespan

Thomas Flatt (S), Elena Pasyukova (S), Amanda Glaser-Schmitt (NS), Eliza Argyridou (NS)

## Size realated traits

### Dry weight 

Hervé Colinet (S), Eran Tauber (S), Banu Sebnem Onder (NS), Katja Hoedjes (NS)

### Wing area

Nico Posnien (S), Nicola Cook (S), Jean-Christophe Billeter (NS), Banu Sebnem Onder (NS)

### Thorax length

Nicola Cook (S), Nico Posnien (S), Iryna Kozeretska (NS), Patrícia Beldade (NS) 


## Temperature related traits

### Chill-coma recovery time

Jorge Vieira (S), Elvira Lafuente (NS), Patricia Gibert (NS)

### Cold-shock mortality

Jorge Vieira (S), Pepi Gonzalez (S), Elvira Lafuente (NS), Iryna Kozeretska (NS) 

### Diapause

Thomas Flatt (S), Alan Bergland (S), Élio (NS) Manolis Lirakis (NS)

### Heat-shock mortality

Jorge Vieira (S), Joyce Kao (S), Amanda Glaser-Schmitt (NS), Eliza Argyridou (NS)


## Other traits 

### Starvation resistance

Elena Pasyukova (S), Pepi Gonzalez (S), Banu Sebnem Onder (NS)

### Pigmentation

Alan Bergland (S), Jessica Abbott (S), Patrícia Beldade (NS), Patricia Gibert (NS)

### Locomotor activity

Eran Tauber (S), Jessica Abbott (S)


# Methods

Contributors and their experimental methods

### Viability
##### Gibert, Grath, Hoedjes, Schmidt, Stamenkovic-Radak and Zwaan Labs
Patricia Gibert, Katja Hoedjes, Paul Schmidt, Marina Stamenkovic-Radak, Bas Zwann, ....

### Developmental time
#### Egg-to-pupa developmental time
##### Schmidt Lab
Paul Schmidt, .... 


#### Egg-to-adult developmental time
##### Gibert, Grath, Hoedjes, Schmidt, Stamenkovic-Radak and Zwaan Labs
Patricia Gibert, Katja Hoedjes, Paul Schmidt, Marina Stamenkovic-Radak, Bas Zwann, .... 

### Dry weight
##### Colinet, Hoedjes and Onder Labs
Herve Colinet, Katja Hoedjes, Banu Onder, .... 

### Thorax length
##### Kozeretska, Posnien, Ritchie and Schmidt Labs

### Wing area
##### Onder, Posnien, Ritchie and Stamenkovic-Radak Labs

### Fecundity
##### Billeter and Fricke Labs

### Lifespan
##### Flatt, Parsch and Pasyukova Labs
Esra Durmaz, Envel Kerdaffrec, TBO, Virginie, John Parsch, Amanda, Eliza, Lena Pasyukova, Alexander 

### Cold-shock mortality
##### Gonzalez, Kozeretska and Vieira Labs

### Chill-coma recovery time
##### Vieira and Mensh Labs

### Heat-shock mortality
##### Parsch and Vieira Labs

### Diapause
##### Bergland, Flatt and Schlotterer Labs

### Circadian eclosion timing
##### Wegener Lab

### Locomotor activity
##### Tauber Lab

### Starvation resistance
##### Gonzalez, Onder and Pasyukova Labs

### Pigmentation
##### Abbott, Gibert and Schmidt



# Results

### Viability
##### Gibert, Grath, Hoedjes, Schmidt, Stamenkovic-Radak and Zwaan Labs

### Developmental time
#### Egg-to-pupa developmental time
##### Schmidt Lab

#### Egg-to-adult developmental time
##### Gibert, Grath, Hoedjes, Schmidt, Stamenkovic-Radak and Zwaan Labs

### Dry weight
##### Colinet, Hoedjes and Onder Labs

### Thorax length
##### Kozeretska, Posnien, Ritchie and Schmidt Labs

### Wing area
##### Onder, Posnien, Ritchie and Stamenkovic-Radak Labs

### Fecundity
##### Billeter and Fricke Labs

### Lifespan
##### Flatt, Parsch and Pasyukova Labs

### Cold-shock mortality
##### Gonzalez, Kozeretska and Vieira Labs

### Chill-coma recovery time
##### Vieira and Mensh Labs

### Heat-shock mortality
##### Parsch and Vieira Labs

### Diapause
##### Bergland, Flatt and Schlotterer Labs

### Circadian eclosion timing
##### Wegener Lab

### Locomotor activity
##### Tauber Lab

### Starvation resistance
##### Gonzalez, Onder and Pasyukova Labs

### Pigmentation
##### Abbott, Gibert and Schmidt


