---
title: "DrosEU_Pheno_environment_PCA"
author: "Ewan Harney"
date: "2024-06-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```

# Climatic Variables

For the analyses of the climatic variables, two time periods were chosen: 30 years (to trace the long-lasting climatic effects) and 30 days (short weather effect, on average over a couple generations for different localities). 14 climatic variables data (see the list below) was downloaded from the NASA database using R package “nasapower” and PCA analysis was done with “FactoMineR” package.  

```{r climate variables, echo=FALSE, message=FALSE}
ClimVarPCA <- read.csv("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/ClimVar4PCA_4Rmd.csv")
knitr::kable(ClimVarPCA, 
             caption = 'List of the climatic variables used in PCA.')
```

Below we show different outputs from the two different versions of the climate PCA (30 days above, 30 years below), summarising some key results.  

## Vector and Coordinate Plots

In the coordinate plots, the populations separate along the two dimensions rather differently for the 30 day and 30 year data sets, possibly because the short-term climate data (30 days) does not include winter (cold season).  

__PCA for 30 days data__  

```{r 30d vector and coord plots, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/Climate_PCA_VecCoord_30d.svg")
```

__PCA for 30 years data__ 

```{r 30y vector and coord plots, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/Climate_PCA_VecCoord_30y.svg")
```

## Scree Plots (eigenvalues)

The first two PCs in both data sets had eigenvalues >1 explaining 94.7% (30 days) and 89.8% (30 years) of the variation. Thus PC1 and PC2 were used in the further analyses.  

__Scree plot for 30 days data__  

```{r 30d scree plot, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/Climate_PCA_Scree_30d.svg")
```

__Scree plot for 30 years data__  

```{r 30y scree plot, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/Climate_PCA_Scree_30y.svg")
```

## Loadings Plots  

There was much overlap in the contributions of different variable to PC1 and PC2. Average temperatures (TS, T2M and T2MWET) all strongly contributed to PC1, while precipitation (PRECTOTCORR) and humidity (RH2M) contributed to PC2. However, frost days was important for PC1 in the 30 year data but not the 30 day data, and while temperature range was important in both data sets, maximum temperatures (TS_MAX and T2M_MAX) were only important in the 30 year data set.

__Loadings for 30 days data__  

```{r 30d loadings, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/Climate_PCA_Loadings_30d.svg")
```

__Loadings for 30 years data__  

```{r 30y loadings, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/Climate_PCA_Loadings_30y.svg")
```

## Correlations with Geographical Coordinates  

In the 30 year dataset PC1 showed strong (< -0.8) negative correlations with latitude and longitude, while in the 30 day dataset PC2 showed a strong (0.8) positive correlation with latitude.

__Correlations for 30 days data__  

```{r 30d correlation plot, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/Climate_PCA_corrplot_30d.svg")
```

__Correlations for 30 years data__  

```{r 30y correlation plot, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/Climate_PCA_corrplot_30y.svg")
```


# Associations between phenotype and climate principal components

Glms were carried out to investigate associations between phenotype principal components (PCs) and climate PCs. For all three different trait combinations (F13, F9 and M9), and for both sets of climate data (30 days and 30 years), separate glms were run considering either phenotype PC 1, 2 or 3 as a response variable, and both climate PCs 1 and 2 as explanatory variables (18 glms total). The summary table below shows the significance of climate variables in explaining each phenotype PC. A permutation based approach was then used to test if correlations between phenotype and climate PCs were greater than expected at random (1000 permutations per phenotype PC:climate PC combination).

## Summary of phenotype x climate principal components
```{r summary of significant PC-PC associations, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/summary_updated_30y-30d_alt-arrange.svg")
```
```
### Signif. codes: ‘***’ 0.001, ‘**’ 0.01, ‘*’ 0.05
```

Results for glms involving phenotype PC1 (all 3 phenotype PCAs and both 30d and 30y climate PCAs) were rather over-dispersed (>2).

## Permutation results for phenotype x climate PCs  

The permutation derived threshold for significance is demarcated by the coloured lines. Purmuted data appear as grey points, and real data appear as a coloured point. Only points that have higher -log~10~ P values (i.e. lower P values) and higher R^2^ values pass this statistical test (P and R^2^ are perfectly correlated). Four correlations between phenotype PC2 and climate PC2 passed the permutation procedure (F13 - 30 day, F13 - 30 year, F9 - 30 day, and M9 - 30 day).  

```{r permutations for PC-PC associations, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/Permutation/all_permutation_results_v3.svg")
```

## Associations between Phenotype PC2 and Climate PC2  

Climatic and phenotypic variables with greater-than-average contributions (loadings) to a given principal component are shown in the accompanying x- and y-axis vector plots, respectively. For males (M9 plots on right hand side), the phenotype PC2 is inverted relative to females (loadings for Life span and Thorax Length differ in direction)

```{r all PC2 versus PC2 plots, echo=FALSE, message=FALSE}
knitr::include_graphics("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/PhenoPC2_ClimPC2_All_Plots.svg")
```

## Summary statistics for glm and permutation results  

```{r read csv files}
## Load data tables for glms

F13_30y_data <- read.csv("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/F9maxP_30y_data.csv")
F13_30d_data <- read.csv("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/F9maxP_30d_data.csv")
F9_30y_data <- read.csv("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/F9_30y_data.csv")
F9_30d_data <- read.csv("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/F9_30d_data.csv")
M9_30y_data <- read.csv("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/M9_30y_data.csv")
M9_30d_data <- read.csv("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/M9_30d_data.csv")

## Load permutation test objects

emp_pr_F13_y <- readRDS("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/emp_pr_F13_y.rds")
emp_pr_F9_y <- readRDS("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/emp_pr_F9_y.rds")
emp_pr_M9_y <- readRDS("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/emp_pr_M9_y.rds")
emp_pr_F13_d <- readRDS("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/emp_pr_F13_d.rds")
emp_pr_F9_d <- readRDS("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/emp_pr_F9_d.rds")
emp_pr_M9_d <- readRDS("EnvironmentalAnalysis/PCA_Pheno_Clim_SI/emp_pr_M9_d.rds")
```

Below are:  
  
1. Summary ANOVA tables for each of the glms considering phenotype PC x climate PC associations, and  
2. Ranked p values for these associations following the permutation procedure.  

## 1. Summary ANOVA tables

### F13: Climate 30 days

```{r F13 30d PC1}
# Phenotype: PC1
F13_PC1_d <- glm(F13_30d_data$PC1_F9maxP ~ F13_30d_data$PC1_clim + F13_30d_data$PC2_clim,data = F13_30d_data)
Anova(F13_PC1_d, test.statistic="F")
```

```{r F13 30d PC2}
# Phenotype: PC2
F13_PC2_d <- glm(F13_30d_data$PC2_F9maxP ~ F13_30d_data$PC1_clim + F13_30d_data$PC2_clim,data = F13_30d_data)
Anova(F13_PC2_d, test.statistic="F")
```

```{r F13 30d PC3}
# Phenotype: PC3
F13_PC3_d <- glm(F13_30d_data$PC3_F9maxP ~ F13_30d_data$PC1_clim + F13_30d_data$PC2_clim,data = F13_30d_data)
Anova(F13_PC3_d, test.statistic="F")
```

### F13: Climate 30 years

```{r F13 30y PC1}
# Phenotype: PC1
F13_PC1_y <- glm(F13_30y_data$PC1_F9maxP ~ F13_30y_data$PC1_clim + F13_30y_data$PC2_clim,data = F13_30y_data)
Anova(F13_PC1_y, test.statistic="F")
```

```{r F13 30y PC2}
# Phenotype: PC2
F13_PC2_y <- glm(F13_30y_data$PC2_F9maxP ~ F13_30y_data$PC1_clim + F13_30y_data$PC2_clim,data = F13_30y_data)
Anova(F13_PC2_y, test.statistic="F")
```

```{r F13 30y PC3}
# Phenotype: PC3
F13_PC3_y <- glm(F13_30y_data$PC3_F9maxP ~ F13_30y_data$PC1_clim + F13_30y_data$PC2_clim,data = F13_30y_data)
Anova(F13_PC3_y, test.statistic="F")
```

### F9: Climate 30 days

```{r F9 30d PC1}
# Phenotype: PC1
F9_PC1_d <- glm(F9_30d_data$PC1_F9 ~ F9_30d_data$PC1_clim + F9_30d_data$PC2_clim,data = F9_30d_data)
Anova(F9_PC1_d, test.statistic="F")
```

```{r F9 30d PC2}
# Phenotype: PC2
F9_PC2_d <- glm(F9_30d_data$PC2_F9 ~ F9_30d_data$PC1_clim + F9_30d_data$PC2_clim,data = F9_30d_data)
Anova(F9_PC2_d, test.statistic="F")
```

```{r F9 30d PC3}
# Phenotype: PC3
F9_PC3_d <- glm(F9_30d_data$PC3_F9 ~ F9_30d_data$PC1_clim + F9_30d_data$PC2_clim,data = F9_30d_data)
Anova(F9_PC3_d, test.statistic="F")
```

### F9: Climate 30 years

```{r F9 30y PC1}
# Phenotype: PC1
F9_PC1_y <- glm(F9_30y_data$PC1_F9 ~ F9_30y_data$PC1_clim + F9_30y_data$PC2_clim,data = F9_30y_data)
Anova(F9_PC1_y, test.statistic="F")
```

```{r F9 30y PC2}
# Phenotype: PC2
F9_PC2_y <- glm(F9_30y_data$PC2_F9 ~ F9_30y_data$PC1_clim + F9_30y_data$PC2_clim,data = F9_30y_data)
Anova(F9_PC2_y, test.statistic="F")
```

```{r F9 30y PC3}
# Phenotype: PC3
F9_PC3_y <- glm(F9_30y_data$PC3_F9 ~ F9_30y_data$PC1_clim + F9_30y_data$PC2_clim,data = F9_30y_data)
Anova(F9_PC3_y, test.statistic="F")
```

### M9: Climate 30 days

```{r M9 30d PC1}
# Phenotype: PC1
M9_PC1_d <- glm(M9_30d_data$PC1_M9 ~ M9_30d_data$PC1_clim + M9_30d_data$PC2_clim,data = M9_30d_data)
Anova(M9_PC1_d, test.statistic="F")
```

```{r M9 30d PC2}
# Phenotype: PC2
M9_PC2_d <- glm(M9_30d_data$PC2_M9 ~ M9_30d_data$PC1_clim + M9_30d_data$PC2_clim,data = M9_30d_data)
Anova(M9_PC2_d, test.statistic="F")
```

```{r M9 30d PC3}
# Phenotype: PC3
M9_PC3_d <- glm(M9_30d_data$PC3_M9 ~ M9_30d_data$PC1_clim + M9_30d_data$PC2_clim,data = M9_30d_data)
Anova(M9_PC3_d, test.statistic="F")
```

### M9: Climate 30 years

```{r M9 30y PC1}
# Phenotype: PC1
M9_PC1_y <- glm(M9_30y_data$PC1_M9 ~ M9_30y_data$PC1_clim + M9_30y_data$PC2_clim,data = M9_30y_data)
Anova(M9_PC1_y, test.statistic="F")
```

```{r M9 30y PC2}
# Phenotype: PC2
M9_PC2_y <- glm(M9_30y_data$PC2_M9 ~ M9_30y_data$PC1_clim + M9_30y_data$PC2_clim,data = M9_30y_data)
Anova(M9_PC2_y, test.statistic="F")
```

```{r M9 30y PC3}
# Phenotype: PC3
M9_PC3_y <- glm(M9_30y_data$PC3_M9 ~ M9_30y_data$PC1_clim + M9_30y_data$PC2_clim,data = M9_30y_data)
Anova(M9_PC3_y, test.statistic="F")
```

## 2. Ranked P values following permutation

The following tables are ranked by pr_p: the probability that the p-value for the observed data is smaller than the permutations. Phenotype PC : climate PC combinations with a pr_p > 0.95 are considered significant.

#### Climate 30 days  

```{r F13 30 day}
# F13: Climate 30 days
emp_pr_F13_d[order(emp_pr_F13_d$pr_p),]
```

```{r F9 30 day}
# F9: Climate 30 days
emp_pr_F9_d[order(emp_pr_F9_d$pr_p),]
```

```{r M9 30 day}
# M9: Climate 30 days
emp_pr_M9_d[order(emp_pr_M9_d$pr_p),]
```

#### Climate 30 years  

```{r F13 30 year}
# F13: Climate 30 years
emp_pr_F13_y[order(emp_pr_F13_y$pr_p),]
```

```{r F9 30 year}
# F9: Climate 30 years
emp_pr_F9_y[order(emp_pr_F9_y$pr_p),]
```

```{r M9 30 year}
# M9: Climate 30 years
emp_pr_M9_y[order(emp_pr_M9_y$pr_p),]
```

