---
title: "DrosEU_phenotype_PCA_diet"
author: "Ewan Harney"
date: "2024-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(psych)
library(knitr)
```

# Effect of diet on PCA

Read in the various PCA objects

```{r read PCA objects}
## Load PCA objects

## Original F13 and F9 PCA objects
pca_F13 <- readRDS("LinearModelsPop/PCA_ControlledDiet/PCA_CD_SI/pca_F13.rds")
pca_F9 <- readRDS("LinearModelsPop/PCA_ControlledDiet/PCA_CD_SI/pca_F9.rds")

## Additional PCA objects for testing effect of diet
pca_M8 <- readRDS("LinearModelsPop/PCA_ControlledDiet/PCA_CD_SI/pca_M8.rds")

pcactl_F13 <- readRDS("LinearModelsPop/PCA_ControlledDiet/PCA_CD_SI/pcactl_F13.rds")
pcactl_F9 <- readRDS("LinearModelsPop/PCA_ControlledDiet/PCA_CD_SI/pcactl_F9.rds")
pcactl_M8 <- readRDS("LinearModelsPop/PCA_ControlledDiet/PCA_CD_SI/pcactl_M8.rds")
```

For F13 and F9 PCAs with all data were compared to PCAs featuring a single estimate for each phenotypic trait value taken from labs in which dietary P:C ratio was similar.

For males, thorax Length (TL) values were not available for the diet controlled PCA, so the PCA with all data was rerun with only the 8 other phenotypic traits.

Key summary results appear below. Loadings for the first five principal components for the all data PCA and the controlled diet data PCA are provided, followed by Tucker's congruence coefficients for these PCs. These coefficients show how well principal componenets (PCs) from the different PCAs correlate with one another (values above 0.85 can be considered high, values above 0.95 very high). Co-ordinate plots (with population confidence ellipses) are then provided for PC1 vs PC2.

### F13 effect of diet 

```{r diet effect for F13 original loadings}
# F13 loadings for all data
pca_F13$var$coord
```

```{r diet effect for F13 diet control loadings}
# F13 loadings for controlled diet data
pcactl_F13$var$coord
```

```{r diet effect for F13 tucker}
# F13 Tucker's congruence values
factor.congruence(pca_F13$var$coord, pcactl_F13$var$coord,2)
```

```{r echo=FALSE}
knitr::include_graphics("LinearModelsPop/PCA_ControlledDiet/PCA_CD_SI/plot_diet_F13.svg")
```

Tucker's congruence coefficient for PC1 is high (0.91), and for PC2 the coefficient is moderate (0.75), with some key correlations between traits appearing in both PCAs (positively correlated loadings for DW, TL and WA in PC1, positively correlated HSM and Via in PC2). The populations thus occupy similar positions within the PC1 vs PC2 co-ordinate space. Although congruence coefficients for PC3, PC4 and PC5 are lower (0.56-0.77), some correlations are maintained (e.g. positively correlated CCRT and Dia in PC3)


### F9 effect of diet 

```{r diet effect for F9 original loadings}
# F9 loadings for all data
pca_F9$var$coord
```

```{r diet effect for F9 diet control loadings}
# F9 loadings for controlled diet data
pcactl_F9$var$coord
```


```{r diet effect for F9 Tucker}
# F9 Tucker's congruence values
factor.congruence(pca_F9$var$coord, pcactl_F9$var$coord,2)
```

```{r echo=FALSE}
knitr::include_graphics("LinearModelsPop/PCA_ControlledDiet/PCA_CD_SI/plot_diet_F9.svg")
```

Tucker's congruence coefficient for PC1 is very high, but coefficients for other traits are lower, with some reordering of PCs 2, 3 and 4), although the ordering of PCs is retained (PC1 best correlates with PC1, PC2 with PC2 etc). As with the F13, PC1 trait correlations are similar (positively correlated loadings for DW, TL and WA in PC1), but correlations are broken or moved (e.g. the negative correlation between SR and TL moves from PC2 in all data to PC3 in the diet controlled PCA). Thus the populations occupy similar positions along PC1 but with differences along PC2 (even accounting for the change in sign).

### M8 effect of diet 

```{r diet effect for M8 original loadings}
# M8 loadings for all data
pca_M8$var$coord
```

```{r diet effect for M8 diet control loadings}
# M8 loadings for controlled diet data
pcactl_M8$var$coord
```

```{r diet effect for M8}
# M8 Tucker's congruence values
factor.congruence(pca_M8$var$coord, pcactl_M8$var$coord,2)
```

```{r echo=FALSE}
knitr::include_graphics("LinearModelsPop/PCA_ControlledDiet/PCA_CD_SI/plot_diet_M8.svg")
```

Tucker's congruence coefficient for PC1 is very high and for PC2 it is high, with many PC1 and PC2 trait correlations similar for the two PCAs, leading to the populations occupying similar positions within the PC1 vs PC2 co-ordinate space.  Congruence coefficients are lower for PC3, PC4 and PC5, although some trait correlations are maintained (e.g. the positive correlation of CCRT and SR along PC3).