---
title: "DrosEU_phenotype_PCA"
author: "Ewan Harney"
output: html_document
date: "2024-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

# PCA

Read in the various PCA objects

```{r read PCA objects}
## Load PCA objects
pca_F13 <- readRDS("PCA/PCA_SI/pca_F13.rds")
pca_F9 <- readRDS("PCA/PCA_SI/pca_F9.rds")
pca_M9 <- readRDS("PCA/PCA_SI/pca_M9.rds")
```

Principal Component Analysis (PCA) was carried out for each sex separately. Initially, a PCA containing all 12 female traits (CCRT, CSM, DT_A, DW, HSM, LS, SR, TL, WA_L, DIA, FEC and PGM_T) plus Viability was considered, termed F13 (Female - 13 traits). Separate female and male PCAs were then carried out for a set of nine traits which had been measured in both sexes (CCRT, CSM, DT_A, DW, HSM, LS, SR, TL, WA_L) termed F9 (Female - traits) and M9 (Male - 9 traits), allowing a direct comparison of the sexes.

Summary results are provided below. Figures show vector plots next to co-ordinate plots (with population confidence ellipses) for PC1 vs PC2 and PC3 vs PC4. Below these are the eigenvalues (the contribution of each principal component to the total variance) and the loadings (how each variable contributes to the first five principal component). Loadings with values > 0.6 or < -0.6 are considered important, while loadings with values > 0.4 or < -0.4 are considered to have moderate contributions.

Considering important and moderate loadings we can identify some general patterns across all three PCA, including:  
  
  * PC1: positive correlations between dry weight, wing area and starvation resistance  
  * PC2: a negative correlation between lifespan and thorax length  
  * PC3: a negative correlation between heat shock mortality and cold shock mortality  

### Female - all traits (F13)

F13 Vector plots (left hand side) and co-ordinate plots (right-hand side) for PC1 vs PC2 (top) and PC3 vs PC4 (bottom). Co-ordinate plots show populations surrounded by 95% confidence ellipses.

```{r echo=FALSE}
knitr::include_graphics("PCA/PCA_SI/plot_F13_1v2.svg")
knitr::include_graphics("PCA/PCA_SI/plot_F13_3v4.svg")
```

```{r eigenvalues for F13}
# F13 eigenvalues
pca_F13$eig
```

```{r loadings for F13}
# F13 loadings
pca_F13$var$coord
```

### Female - 9 traits (F9)

F9 Vector plots (left hand side) and co-ordinate plots (right-hand side) for PC1 vs PC2 (top) and PC3 vs PC4 (bottom). Co-ordinate plots show populations surrounded by 95% confidence ellipses.

```{r echo=FALSE}
knitr::include_graphics("PCA/PCA_SI/plot_F9_1v2.svg")
knitr::include_graphics("PCA/PCA_SI/plot_F9_3v4.svg")
```


```{r eigenvalues for F9}
# F9 eigenvalues
pca_F9$eig
```

```{r loadings for F9}
# F9 loadings
pca_F9$var$coord
```


### Male - 9 traits (M9)

M9 Vector plots (left hand side) and co-ordinate plots (right-hand side) for PC1 vs PC2 (top) and PC3 vs PC4 (bottom). Co-ordinate plots show populations surrounded by 95% confidence ellipses. PC2 

```{r echo=FALSE}
knitr::include_graphics("PCA/PCA_SI/plot_M9_1v2.svg")
knitr::include_graphics("PCA/PCA_SI/plot_M9_3v4.svg")
```

```{r eigenvalues for M9}
# M9 eigenvalues
pca_M9$eig
```

```{r loadings for M9}
# M9 loadings
pca_M9$var$coord
```
